\documentclass[preprint,12pt,authoryear]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{setspace}

\journal{Accident Analysis \& Prevention}
\doublespacing
\modulolinenumbers[1]
\linenumbers

\begin{document}

\begin{frontmatter}

\title{Evidence-Bounded Driver Process-Model Replay for Automated-Driving Incident Analysis}

\author[aff1]{First Author}
\author[aff1]{Second Author}
\author[aff1]{Corresponding Author\corref{cor1}}
\ead{corresponding.author@example.com}
\cortext[cor1]{Corresponding author.}
\affiliation[aff1]{organization={Department or Laboratory Name},
  addressline={University or Institute Name},
  city={City},
  country={Country}}

\begin{abstract}
Public automated-driving crash and disengagement reports provide valuable real-world evidence, but they often omit the driver-visible HMI, driver state, transition timing, and internal ADS information needed for strong driver-centered causal claims. This paper proposes an STPA-HF-constrained Evidence-Bounded Driver Process-Model Replay framework that transforms sparse incident text into auditable tabletop replay packages. The framework represents driver process-model variables, process-model update sources, candidate driver control actions, unsafe-control-action pathways, outcome compatibility, blocked claims, and minimum reporting/logging requirements. The study evaluates whether this replay layer can generate evidence-grounded review artifacts, suppress outcome-only overreach relative to direct LLM and generic chain-of-thought baselines, align with expert review of ranked pathways, and respond consistently to HMI/logging counterfactual injections. A richer-evidence case study is reserved as an evidence-density stress test.
\end{abstract}

\begin{keyword}
Automated driving \sep STPA-HF \sep driver process model \sep tabletop replay \sep large language models \sep human factors
\end{keyword}

\end{frontmatter}

\input{introduction}
\input{related_work}
\input{method_spine}
\input{experiment_plan}

\section*{Declaration of competing interest}
Declarations of interest: none.

\section*{Data availability}
The public incident reports and derived functional cases used in this study will be described in the final manuscript. Code and processed artifacts will be made available subject to data-source licensing and anonymization constraints.

\bibliographystyle{elsarticle-harv}
\bibliography{references}

\end{document}
